---
title: "Replicating Clark & McCracken (2001)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replicating Clark & McCracken (2001)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 4
)
```

This article reproduces the Phillips-curve forecasting illustration
of Clark and McCracken (2001, Section 5) on the bundled `cm2001`
dataset. The benchmark forecasts US unemployment from its own lags
(univariate AR); the alternative adds lagged inflation. `enc_new()`
tests whether the AR forecasts *encompass* those of the augmented
ARX model — i.e., whether inflation adds any predictive content
beyond what lagged unemployment already contains.

```{r setup, message = FALSE}
library(forecastdom)
data(cm2001)

# CM (2001) sample: 1957:01-1997:08.
cm <- subset(cm2001,
             date >= as.Date("1957-01-01") &
             date <= as.Date("1997-08-01"))
nrow(cm)
```

## Recursive forecasts

Both models are estimated by OLS on an expanding window with `p`
lags of each variable. At each origin we form a one-step-ahead
forecast, then roll forward by one month.

```{r helper}
make_lags <- function(z, p) {
  n <- length(z)
  sapply(seq_len(p), function(k) c(rep(NA, k), z[seq_len(n - k)]))
}

recursive_arx <- function(y, x, p, R) {
  n <- length(y)
  YL <- make_lags(y, p)
  XL <- make_lags(x, p)
  target <- y[(p + 1):n]
  YL <- YL[(p + 1):n, , drop = FALSE]
  XL <- XL[(p + 1):n, , drop = FALSE]
  P <- length(target) - R
  e_ar <- e_arx <- numeric(P)
  for (j in seq_len(P)) {
    idx     <- seq_len(R + j - 1)
    fit_ar  <- lm.fit(cbind(1, YL[idx, ]),         target[idx])
    fit_arx <- lm.fit(cbind(1, YL[idx, ], XL[idx, ]), target[idx])
    pred_ar  <- sum(coef(fit_ar)  * c(1, YL[R + j, ]))
    pred_arx <- sum(coef(fit_arx) * c(1, YL[R + j, ], XL[R + j, ]))
    e_ar[j]  <- target[R + j] - pred_ar
    e_arx[j] <- target[R + j] - pred_arx
  }
  list(e_ar = e_ar, e_arx = e_arx)
}
```

## ENC-NEW across lag orders

CM (2001) consider lag orders `p = 1, 3, 6, 12`. The initial
estimation window is `R = 120` months (10 years); the remaining
observations form the out-of-sample period.

```{r encnew-table}
R <- 120L
rows <- lapply(c(1L, 3L, 6L, 12L), function(p) {
  fc  <- recursive_arx(cm$unrate, cm$infl, p = p, R = R)
  enc <- enc_new(fc$e_ar, fc$e_arx)
  msfe1 <- mean(fc$e_ar^2)
  msfe2 <- mean(fc$e_arx^2)
  data.frame(p        = p,
             n_oos    = length(fc$e_ar),
             pi_ratio = round(length(fc$e_ar) / R, 2),
             MSFE_AR  = msfe1,
             MSFE_ARX = msfe2,
             R2OS_pct = 100 * (1 - msfe2 / msfe1),
             ENC_NEW  = unname(enc$statistic))
})
tab <- do.call(rbind, rows)
knitr::kable(
  tab, digits = 3, row.names = FALSE,
  col.names = c("$p$", "$T_{OOS}$", "$\\pi = P/R$",
                "$MSFE_{AR}$", "$MSFE_{ARX}$",
                "$R^2_{OS}$ (%)", "ENC-NEW"))
```

For every lag order considered, the ARX model achieves a lower MSFE
than the pure AR — that is, knowing past inflation reduces
out-of-sample unemployment-forecast error. ENC-NEW values are large
relative to the Clark-McCracken (2001) Table 2 asymptotic 5%
critical values, which are tabulated in `k2` (number of extra
regressors = `p`) and `π = P/R`. For `k2 = 6`, `π ≈ 2`, the 5% CV is
≈ 8.6; for `k2 = 12`, `π ≈ 2`, the 5% CV is ≈ 12.0. The realised
statistics in the table exceed these benchmarks decisively in every
specification, so the AR forecasts do **not** encompass the ARX —
inflation carries real predictive content for unemployment.

## Comparison with Clark-West MSFE-adjusted

`enc_new()` answers an encompassing question; `cw_test()` answers
the closely related "equal MSFE" question with a t-statistic that
has a standard normal asymptotic distribution. The two tests
typically point the same way; reporting both is informative.

```{r cw-comparison}
rows2 <- lapply(c(1L, 3L, 6L, 12L), function(p) {
  fc <- recursive_arx(cm$unrate, cm$infl, p = p, R = R)
  f1 <- cm$unrate[(p + 1 + R):nrow(cm)] - fc$e_ar
  f2 <- cm$unrate[(p + 1 + R):nrow(cm)] - fc$e_arx
  cw <- cw_test(fc$e_ar, fc$e_arx, f1, f2)
  data.frame(p         = p,
             CW_stat   = unname(cw$statistic),
             CW_pvalue = unname(cw$pvalue))
})
knitr::kable(
  do.call(rbind, rows2), digits = 3, row.names = FALSE,
  col.names = c("$p$", "CW stat", "CW $p$-value"))
```

CW p-values are far below 5% for every lag order, agreeing with the
ENC-NEW decision: the ARX model has significantly lower out-of-sample
MSFE than the pure AR.

## References

- Clark, T. E. and McCracken, M. W. (2001). Tests of equal forecast
  accuracy and encompassing for nested models. *Journal of
  Econometrics*, 105(1), 85-110.
- Clark, T. E. and West, K. D. (2007). Approximately normal tests
  for equal predictive accuracy in nested models. *Journal of
  Econometrics*, 138(1), 291-311.
