---
title: "Clark-West Tests on Equity Premium Forecasts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Clark-West Tests on Equity Premium Forecasts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 4
)
```

This article applies the Clark-West (2007) MSFE-adjusted test --
exposed by **forecastdom** as `cw_test()` -- to two classic equity
premium forecasting datasets bundled with the package:

* `rz2013` -- the 14 Goyal-Welch macroeconomic predictors used in
  Rapach and Zhou (2013), *Forecasting Stock Returns*, Handbook of
  Economic Forecasting, Vol. 2A, Chapter 6.
* `nrtz2014` -- the 14 binary technical indicators (moving averages,
  momentum, on-balance volume) of Neely, Rapach, Tu, and Zhou (2014),
  *Forecasting the Equity Risk Premium*, *Management Science*.

The benchmark in both exercises is the prevailing historical mean of
the equity premium -- the standard "no predictability" null in this
literature -- and the alternatives are bivariate predictive regressions
on each single predictor.

```{r setup, message = FALSE}
library(forecastdom)
library(ggplot2)

data(rz2013)
data(nrtz2014)
```

## A reusable forecasting helper

The same recursive (expanding-window) loop is used for every
predictor. At each forecast date \(t+1\), the historical-mean
benchmark uses \(\bar y_{1:t}\); the alternative regresses
\(y_{2:t}\) on \(x_{1:(t-1)}\) and forecasts using \(x_t\).

```{r helper}
recursive_forecasts <- function(y, x, R) {
  T_ <- length(y)
  P  <- T_ - R
  e1 <- e2 <- f1 <- f2 <- numeric(P)

  for (j in seq_len(P)) {
    # estimation sample: rows 1..(R+j-1); forecast target: row R+j
    train_y <- y[1:(R + j - 1)]
    train_x <- x[1:(R + j - 1)]

    # benchmark: prevailing mean of y observed up to t
    f1[j] <- mean(train_y)

    # alternative: y_t = a + b * x_{t-1} + e
    fit <- lm(yt ~ xlag, data = data.frame(
      yt   = train_y[-1],
      xlag = train_x[-length(train_x)]
    ))
    f2[j] <- as.numeric(stats::predict(
      fit, newdata = data.frame(xlag = train_x[length(train_x)])
    ))

    e1[j] <- y[R + j] - f1[j]
    e2[j] <- y[R + j] - f2[j]
  }

  list(e1 = e1, e2 = e2, f1 = f1, f2 = f2, P = P)
}
```

## Macro predictors (Rapach & Zhou, 2013)

We use the first 241 monthly observations (1926-12 through 1946-12) as
the initial estimation window and evaluate the next 768 months
(1947-01 through 2010-12).

```{r rz-setup}
y_rz   <- rz2013$eq_prem
preds_rz <- setdiff(names(rz2013), c("date", "eq_prem"))
R_rz   <- 241L

cw_rz <- lapply(preds_rz, function(p) {
  fc <- recursive_forecasts(y_rz, rz2013[[p]], R = R_rz)
  res <- cw_test(fc$e1, fc$e2, fc$f1, fc$f2)
  data.frame(
    predictor = p,
    R2OS_pct  = unname(res$r2os),
    CW_stat   = unname(res$statistic),
    p_value   = unname(res$pvalue)
  )
})
cw_rz <- do.call(rbind, cw_rz)
cw_rz <- cw_rz[order(-cw_rz$CW_stat), ]

knitr::kable(cw_rz, digits = 3, row.names = FALSE,
             caption = "Clark-West test, macro predictors vs. historical mean (rz2013)")
```

A positive `R2OS_pct` indicates the alternative model has a lower
out-of-sample MSFE than the historical-mean benchmark; the
Clark-West *t*-statistic is compared to a one-sided standard normal
under \(H_0:\, \text{R}^2_{OS} \le 0\).

## Technical indicators (Neely et al., 2014)

The technical indicators are binary signals constructed from S&P 500
prices and volume. We use the first 181 observations (1950-12 through
1965-12) as the initial estimation window and evaluate the next 552
months (1966-01 through 2011-12), matching the paper.

```{r nrtz-setup}
y_nr   <- nrtz2014$eq_prem
preds_nr <- setdiff(names(nrtz2014), c("date", "eq_prem"))
R_nr   <- 181L

cw_nr <- lapply(preds_nr, function(p) {
  fc <- recursive_forecasts(y_nr, nrtz2014[[p]], R = R_nr)
  res <- cw_test(fc$e1, fc$e2, fc$f1, fc$f2)
  data.frame(
    predictor = p,
    R2OS_pct  = unname(res$r2os),
    CW_stat   = unname(res$statistic),
    p_value   = unname(res$pvalue)
  )
})
cw_nr <- do.call(rbind, cw_nr)
cw_nr <- cw_nr[order(-cw_nr$CW_stat), ]

knitr::kable(cw_nr, digits = 3, row.names = FALSE,
             caption = "Clark-West test, technical indicators vs. historical mean (nrtz2014)")
```

Most technical indicators deliver positive R²OS and *Clark-West* statistics that
are significant at conventional levels, consistent with the headline
finding of Neely et al. (2014): simple price-based and volume-based
signals contain economically meaningful information for the equity
premium.

## A single deep-dive: MA(2, 12)

```{r ma-print}
fc_ma <- recursive_forecasts(nrtz2014$eq_prem, nrtz2014$MA_2_12, R = R_nr)
cw_test(fc_ma$e1, fc_ma$e2, fc_ma$f1, fc_ma$f2)
```

The cumulative squared-error differential (Goyal-Welch, 2008) tracks
*when* the predictor outperforms the historical mean. An upward-
sloping line means the technical indicator is doing better; the steep
gains tend to cluster around recessions, exactly the regime
documented by Neely et al. (2014).

```{r ma-plot}
oos_dates <- tail(nrtz2014$date, fc_ma$P)
loss_diff <- fc_ma$e1^2 - fc_ma$e2^2

ggplot(data.frame(date = oos_dates, cum = cumsum(loss_diff)),
       aes(date, cum)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line(colour = "#47A5C5", linewidth = 0.8) +
  labs(x = NULL,
       y = "Cumulative SE diff (mean - MA(2,12))",
       title = "Cumulative squared-error differential, MA(2,12) vs. historical mean",
       subtitle = "Above zero means the technical indicator is winning") +
  theme_minimal(base_size = 11)
```

## Takeaways

* `cw_test()` is a one-line drop-in for any expanding-window equity
  premium forecasting exercise: feed it the two error series and the
  two forecast series, and read off the MSFE-adjusted statistic, the
  one-sided *p*-value, and \(R^2_{OS}\).
* For the Goyal-Welch macro predictors (`rz2013`), out-of-sample gains
  over the historical mean are economically modest and only a handful
  of predictors are significant -- the well-known "Goyal-Welch
  puzzle".
* For the technical indicators (`nrtz2014`), most signals deliver
  positive \(R^2_{OS}\) with significant Clark-West statistics, in
  line with the central message of Neely et al. (2014).

## References

- Clark, T. E. and West, K. D. (2007). Approximately normal tests for
  equal predictive accuracy in nested models. *Journal of
  Econometrics*, 138(1), 291-311.
- Goyal, A. and Welch, I. (2008). A comprehensive look at the
  empirical performance of equity premium prediction. *Review of
  Financial Studies*, 21(4), 1455-1508.
- Neely, C. J., Rapach, D. E., Tu, J., and Zhou, G. (2014). Forecasting
  the equity risk premium: The role of technical indicators.
  *Management Science*, 60(7), 1772-1791.
- Rapach, D. E. and Zhou, G. (2013). Forecasting stock returns. In
  G. Elliott and A. Timmermann (Eds.), *Handbook of Economic
  Forecasting*, Vol. 2A, pp. 328-383. Elsevier.
