---
title: "Replicating Rapach & Zhou (2013) and Neely et al. (2014)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replicating Rapach & Zhou (2013) and Neely et al. (2014)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 4
)
```

This article applies `cw_test()` to two bundled equity premium
datasets:

* `rz2013` -- 14 Goyal-Welch macroeconomic predictors used in
  Rapach and Zhou (2013).
* `nrtz2014` -- 14 binary technical indicators used in Neely,
  Rapach, Tu and Zhou (2014).

In both cases the benchmark is the prevailing historical mean and
the alternative is a bivariate predictive regression on a single
predictor.

```{r setup, message = FALSE}
library(forecastdom)
data(rz2013); data(nrtz2014)
```

## Helper

```{r helper}
recursive_forecasts <- function(y, x, R) {
  P <- length(y) - R
  e1 <- e2 <- f1 <- f2 <- numeric(P)
  for (j in seq_len(P)) {
    ty <- y[1:(R + j - 1)]
    tx <- x[1:(R + j - 1)]
    f1[j] <- mean(ty)
    fit   <- lm(yt ~ xlag, data = data.frame(yt = ty[-1], xlag = tx[-length(tx)]))
    f2[j] <- as.numeric(predict(fit, newdata = data.frame(xlag = tx[length(tx)])))
    e1[j] <- y[R + j] - f1[j]
    e2[j] <- y[R + j] - f2[j]
  }
  list(e1 = e1, e2 = e2, f1 = f1, f2 = f2)
}

run_cw <- function(data, predictors, R) {
  do.call(rbind, lapply(predictors, function(p) {
    fc  <- recursive_forecasts(data$eq_prem, data[[p]], R = R)
    res <- cw_test(fc$e1, fc$e2, fc$f1, fc$f2)
    data.frame(predictor = p,
               R2OS_pct  = unname(res$r2os),
               CW_stat   = unname(res$statistic),
               p_value   = unname(res$pvalue))
  }))
}
```

## Macro predictors (Rapach & Zhou, 2013)

Initial estimation window of 241 months (1926-12 to 1946-12); OOS
period 1947-01 to 2010-12.

```{r rz}
preds_rz <- c("DP", "EP", "NTIS", "TBL", "INFL_lag")
knitr::kable(run_cw(rz2013, preds_rz, R = 241), digits = 3, row.names = FALSE)
```

## Technical indicators (Neely et al., 2014)

Initial window of 181 months (1950-12 to 1965-12); OOS period
1966-01 to 2011-12.

```{r nrtz}
preds_nr <- c("MA_1_9", "MA_2_12", "MOM_9", "MOM_12", "VOL_2_12")
knitr::kable(run_cw(nrtz2014, preds_nr, R = 181), digits = 3, row.names = FALSE)
```

## Takeaways

* For Goyal-Welch macro predictors, out-of-sample gains over the
  historical mean are economically small -- the well-known
  Goyal-Welch puzzle.
* Technical indicators deliver positive \(R^2_{OS}\) with
  Clark-West statistics significant at conventional levels --
  the central finding of Neely et al. (2014).

## References

- Clark, T. E. and West, K. D. (2007). Approximately normal tests
  for equal predictive accuracy in nested models. *Journal of
  Econometrics*, 138(1), 291-311.
- Neely, C. J., Rapach, D. E., Tu, J. and Zhou, G. (2014).
  Forecasting the equity risk premium: The role of technical
  indicators. *Management Science*, 60(7), 1772-1791.
- Rapach, D. E. and Zhou, G. (2013). Forecasting stock returns.
  In G. Elliott and A. Timmermann (Eds.), *Handbook of Economic
  Forecasting*, Vol. 2A, pp. 328-383. Elsevier.
