---
title: "Get Started with forecastdom"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get Started with forecastdom}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4
)
```

## Overview

**forecastdom** provides a comprehensive suite of forecast evaluation tests organized around the taxonomy proposed by Li, Liao, and Quaedvlieg (2022):

|  | **Equal** | **Superior** |
|---|---|---|
| **Unconditional** | `dm_test()` | `spa_test()` |
| **Conditional** | `gw_test()` | `cspa_test()` |

Along with complementary tests for nested models (`cw_test()`, `enc_new()`), predictive regressions (`ivx_wald()`), and parameter instability (`qll_hat()`).

```{r setup}
library(forecastdom)
set.seed(42)
```

## Pairwise Forecast Comparison

### Diebold-Mariano Test

The simplest starting point: compare two sets of forecast errors.

```{r dm}
e1 <- rnorm(200)
e2 <- rnorm(200, mean = 0.15)
dm_test(e1, e2)
```

The `correction = TRUE` default applies the Harvey, Leybourne, and Newbold (1997) small-sample correction with t-distribution critical values.

### Clark-West Test

For nested models, the standard DM test is undersized. The Clark-West (2007) MSFE-adjusted test corrects for this:

```{r cw}
actual <- rnorm(200)
f1 <- actual + rnorm(200, sd = 0.5)   # benchmark (restricted)
f2 <- actual + rnorm(200, sd = 0.4)   # alternative (unrestricted)
cw_test(actual - f1, actual - f2, f1, f2)
```

### Giacomini-White Test

Tests whether two methods have *conditionally* equal predictive ability, using instruments from the information set:

```{r gw}
gw_test(e1, e2)
```

## Multiple Forecast Comparison

### Hansen's SPA Test

Tests whether a benchmark is unconditionally superior to all competitors simultaneously:

```{r spa}
sim <- do_sim(J = 3, n = 250, a = 1, c = 0, rho_u = 0.4)
spa_test(sim$Y, level = 0.05, B = 1000)
```

### CSPA Test

The main contribution of Li, Liao, and Quaedvlieg (2022). Tests whether the benchmark's conditional expected loss is no more than that of any competitor, *uniformly* across all conditioning states:

```{r cspa-null}
# Under the null (a = 1): benchmark weakly dominates
sim_null <- do_sim(J = 3, n = 500, a = 1, c = 0, rho_u = 0.4)
cspa_test(sim_null$Y, sim_null$X, level = 0.05, trim = 2)
```

```{r cspa-alt}
# Under the alternative (a = 1.5): a competitor outperforms in some states
sim_alt <- do_sim(J = 3, n = 500, a = 1.5, c = 0, rho_u = 0.4)
result <- cspa_test(sim_alt$Y, sim_alt$X, level = 0.05, trim = 2)
result
```

### Visualization

The `cspa_test_plot()` function displays the estimated conditional mean functions $\hat{h}_j(x)$, their lower envelope, and the confidence bound:

```{r cspa-plot}
cspa_test_plot(result)
```

### CSMS: Confidence Set for the Most Superior

When there is no *a priori* benchmark, the CSMS inverts the CSPA test for each candidate and collects those that are not rejected:

```{r csms, eval = FALSE}
csms(losses, X, level = 0.05, trim = 2,
     method_names = c("AR1", "HAR", "HARQ", "LASSO"))
```

## Predictive Regressions

### IVX-Wald Test

Robust test for return predictability with potentially persistent predictors:

```{r ivx}
n <- 300
x <- cumsum(rnorm(n))
y <- 0.02 * x + rnorm(n)
ivx_wald(y, as.matrix(x), K = 1, M_n = floor(n^(1/3)))
```

### Elliott-Muller Test

Tests the null hypothesis of constant regression coefficients against the alternative of general time variation:

```{r qll}
X <- matrix(rnorm(n * 2), n, 2)
y2 <- X %*% c(0.5, -0.3) + rnorm(n)
qll_hat(y2, X, L = floor(n^(1/3)))
```

## References

- Clark, T.E. and McCracken, M.W. (2001). Tests of Equal Forecast Accuracy and Encompassing for Nested Models. *Journal of Econometrics*, 105(1), 85-110.
- Clark, T.E. and West, K.D. (2007). Approximately Normal Tests for Equal Predictive Accuracy in Nested Models. *Journal of Econometrics*, 138(1), 291-311.
- Diebold, F.X. and Mariano, R.S. (1995). Comparing Predictive Accuracy. *Journal of Business & Economic Statistics*, 13(3), 253-263.
- Elliott, G. and Muller, U.K. (2006). Efficient Tests for General Persistent Time Variation in Regression Coefficients. *Review of Economic Studies*, 73(4), 907-940.
- Giacomini, R. and White, H. (2006). Tests of Conditional Predictive Ability. *Econometrica*, 74(6), 1545-1578.
- Hansen, P.R. (2005). A Test for Superior Predictive Ability. *Journal of Business & Economic Statistics*, 23(4), 365-380.
- Harvey, D., Leybourne, S., and Newbold, P. (1997). Testing the Equality of Prediction Mean Squared Errors. *International Journal of Forecasting*, 13(2), 281-291.
- Kostakis, A., Magdalinos, T., and Stamatogiannis, M.P. (2015). Robust Econometric Inference for Stock Return Predictability. *Review of Financial Studies*, 28(5), 1506-1553.
- Li, J., Liao, Z., and Quaedvlieg, R. (2022). Conditional Superior Predictive Ability. *Review of Economic Studies*, 89(2), 843-875.
