---
title: "Replicating Giacomini & White (2006)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replicating Giacomini & White (2006)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 4
)
```

This article applies `gw_test()` — Giacomini and White (2006)
conditional equal predictive ability (CEPA) test — to the Survey of
Professional Forecasters' mean CPI inflation forecasts. The
benchmark is a naive random walk in inflation (next quarter's
forecast = this quarter's realised inflation); the alternative is
the SPF mean at horizons $h = 0, 1, 2, 3, 4$ quarters. The
bundled `gw2006` dataset extends GW's quarterly sample to the
present using the same Philadelphia Fed SPF source.

```{r setup, message = FALSE}
library(forecastdom)
data(gw2006)
str(gw2006)
```

## CEPA across forecast horizons

```{r table-horizons}
run_h <- function(h) {
  spf_col <- paste0("spf_h", h)
  ok <- complete.cases(gw2006[, c("infl", spf_col, "infl_lag")])
  d  <- gw2006[ok, ]
  e_spf <- d$infl - d[[spf_col]]    # SPF errors
  e_rw  <- d$infl - d$infl_lag      # random-walk errors
  r <- gw_test(e_spf, e_rw)         # default: constant + lagged loss diff
  data.frame(h = h, n = nrow(d),
             mse_spf = mean(e_spf^2),
             mse_rw  = mean(e_rw^2),
             wald    = unname(r$statistic),
             pvalue  = unname(r$pvalue),
             reject  = unname(r$pvalue) < 0.05)
}
tab <- do.call(rbind, lapply(0:4, run_h))
knitr::kable(
  tab, digits = 3, row.names = FALSE,
  col.names = c("$h$", "$n$",
                "$MSE_{SPF}$", "$MSE_{RW}$",
                "Wald", "$p$-value", "Reject"))
```

At every horizon SPF has lower mean squared error than the
random-walk benchmark, yet the GW test fails to reject conditional
equal predictive ability at the 5% level. The pattern is consistent
with Atkeson and Ohanian (2001): once the autocorrelation of the
loss differential is accounted for, sophisticated inflation
forecasts are hard to distinguish from a "no-change" benchmark in a
formal test.

## Choice of instruments

`gw_test()` defaults to two instruments: a constant and the lagged
loss differential. Different instruments target different features
of the conditioning information set. Below we replace the lag with
the lagged inflation level — a regime indicator that captures
high/low inflation environments — at horizon $h = 1$.

```{r instrument-comparison}
h <- 1L
ok <- complete.cases(gw2006[, c("infl", paste0("spf_h", h), "infl_lag")])
d  <- gw2006[ok, ]
e_spf <- d$infl - d[[paste0("spf_h", h)]]
e_rw  <- d$infl - d$infl_lag

# Default: constant + lagged loss differential
r_default <- gw_test(e_spf, e_rw)

# Custom: constant + lagged inflation level (padded with NA at t = 1
# so that nrow(W) matches length(e1)).
W_infl <- cbind(1, c(NA, head(d$infl_lag, -1)))
keep   <- !is.na(W_infl[, 2])
r_infl <- gw_test(e_spf[keep], e_rw[keep],
                  instruments = W_infl[keep, , drop = FALSE])

# Absolute-error loss instead of squared-error
r_ae <- gw_test(e_spf, e_rw, loss = "AE")

tab2 <- data.frame(
  spec = c("Default (const + lagged Δloss)",
           "Const + lagged inflation",
           "Absolute-error loss"),
  wald   = c(r_default$statistic, r_infl$statistic, r_ae$statistic),
  df     = c(r_default$df,        r_infl$df,        r_ae$df),
  pvalue = c(r_default$pvalue,    r_infl$pvalue,    r_ae$pvalue)
)
knitr::kable(
  tab2, digits = 3, row.names = FALSE,
  col.names = c("Specification", "Wald", "df", "$p$-value"))
```

The decision is robust across all three specifications: SPF and
the random walk cannot be statistically distinguished at the 5%
level regardless of which conditioning instruments or loss function
we choose.

## Note on the original paper

Giacomini and White (2006, Section 4) compare SPF nowcasts to the
*Greenbook* — the Federal Reserve staff's internal inflation
forecast — using a richer instrument set. They find that the test
*does* reject equal conditional predictive ability between the two
*sophisticated* forecasts in some conditioning states. Replicating
that result requires Greenbook data with its five-year embargo;
this article instead demonstrates the test mechanics against the
simpler random-walk benchmark, which is freely available.

## References

- Atkeson, A. and Ohanian, L. E. (2001). Are Phillips curves useful
  for forecasting inflation? *Federal Reserve Bank of Minneapolis
  Quarterly Review*, 25(1), 2-11.
- Giacomini, R. and White, H. (2006). Tests of conditional
  predictive ability. *Econometrica*, 74(6), 1545-1578.
