---
title: "Replicating Hansen & Lunde (2005)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replicating Hansen & Lunde (2005)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 4
)
```

This article reproduces the central question of Hansen and Lunde
(2005, *JAE*) — *Does anything beat a GARCH(1,1)?* — using
`spa_test()` on the bundled `hl2005` dataset. The benchmark is
GARCH(1,1) with constant mean and Gaussian errors; the alternatives
are 329 other GARCH-family specifications. The realized-variance
proxy choice turns out to matter for the decision.

```{r setup, message = FALSE}
library(forecastdom)
data(hl2005)

n <- length(hl2005$date)
J <- ncol(hl2005$forecasts)
sprintf("Sample: %s to %s (%d trading days), %d forecast models.",
        hl2005$date[1], hl2005$date[n], n, J)
```

## Setup

Squared-error loss `(f - y)^2` against each RV proxy. The
loss differential matrix `Y` is `competitor_loss − benchmark_loss`,
so positive values mean GARCH(1,1) wins.

```{r setup-losses}
b <- hl2005$garch11_idx

build_Y <- function(rv) {
  L <- (hl2005$forecasts - rv)^2
  L[, -b] - L[, b]
}

Y <- build_Y(hl2005$rv)            # primary RV proxy (5-min linear)
dim(Y)

cat("Competitors with lower MSE than GARCH(1,1):",
    sum(colMeans(Y) < 0), "of", ncol(Y), "\n")
```

More than half of the 329 alternatives beat GARCH(1,1) on average
loss — the unconditional ranking does not single out GARCH(1,1).
The question is whether any of them does so by a *statistically
significant* margin after correcting for multiple testing. That is
exactly what `spa_test()` answers.

## SPA test against the primary RV proxy

```{r primary-spa}
set.seed(20260512)
r <- spa_test(Y, level = 0.05, B = 5000L, q = 0.25)
r
```

T_SPA is large but the bootstrap p-value sits just above 5%.
Conclusion: at the 5% level, with the 5-minute linear-interpolation
proxy, *nothing significantly beats GARCH(1,1)* — Hansen and Lunde's
headline finding.

## Robustness across realised-variance proxies

The dataset ships with eight different RV proxies, from the very
noisy squared close-to-close return to fine 1-minute sampled
estimators. Re-running the SPA test across all eight produces the
paper's central robustness table.

```{r proxy-grid}
proxies <- colnames(hl2005$rv_proxies)

tab <- do.call(rbind, lapply(proxies, function(p) {
  set.seed(20260512)
  r <- spa_test(build_Y(hl2005$rv_proxies[, p]),
                level = 0.05, B = 5000L, q = 0.25)
  data.frame(proxy = p,
             T_SPA   = unname(r$statistic),
             pvalue  = unname(r$pvalue),
             reject  = unname(r$reject),
             n_beat  = sum(colMeans(build_Y(hl2005$rv_proxies[, p])) < 0))
}))

knitr::kable(tab, digits = 3, row.names = FALSE)
```

The decision flips with proxy quality:

- **Noisy proxy** (`sq_ccr`, squared close-to-close returns): T_SPA
  much smaller and p-value far from rejection. With a noisy target
  no model can be confidently ranked against any other.
- **Coarse intraday proxies** (5-min linear/previous-tick): p-values
  hover near 0.07-0.08 — just *failing* to reject GARCH(1,1).
- **Fine intraday proxies** (1-min linear/previous-tick, Fourier,
  Spline-250): p-values fall below 5% and the SPA *does* reject —
  with a sufficiently accurate volatility proxy, some competitor
  models can be shown to beat GARCH(1,1).

This is precisely Hansen and Lunde's nuanced answer: GARCH(1,1) is
hard to beat in any *concrete* unconditional comparison, but the
decision is sensitive to how cleanly we measure realised volatility.

## Top 10 alternatives by mean loss (5-min proxy)

```{r top10}
d_bar <- colMeans(Y)
top10_idx <- order(d_bar)[1:10]
data.frame(rank = 1:10,
           competitor_col = (1:J)[-b][top10_idx],
           mean_loss_diff = round(d_bar[top10_idx], 3))
```

`mean_loss_diff` is the average of `L_competitor − L_GARCH(1,1)`;
negative values mean the competitor has lower mean MSE than
GARCH(1,1). Columns 261-265 correspond to the EGARCH family with
constant mean and t-distributed errors (per the README's grouping
of the 330 specifications) — the most consistent winners across
proxies.

## References

- Hansen, P. R. (2005). A test for superior predictive ability.
  *Journal of Business & Economic Statistics*, 23(4), 365-380.
- Hansen, P. R. and Lunde, A. (2005). A forecast comparison of
  volatility models: does anything beat a GARCH(1,1)?
  *Journal of Applied Econometrics*, 20(7), 873-889.
