---
title: "Replicating Li et al. (2022)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replicating Li et al. (2022)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 4
)
```

This article reproduces the volatility-forecasting application of Li,
Liao and Quaedvlieg (2022, *RFS*), Section 4: Figure 2
(one-vs-one CSPA on JNJ), Figure 3 (one-vs-all CSPA on JNJ), and
Table 4 Panel A (cross-stock rejection counts). All numbers and
plots are generated from the bundled `llq2022_jnj`, `llq2022` and
`llq2022_uv_cspa` datasets.

```{r setup, message = FALSE}
library(forecastdom)
data(llq2022_jnj)      # JNJ realized variance + 6 forecasts + lagged VIX
data(llq2022)          # S&P 500 counterpart
data(llq2022_uv_cspa)  # pre-computed cross-stock counts

models <- c("AR1", "AR22", "AR22_Lasso", "HAR", "HARQ", "ARFIMA")
qlike  <- function(f, y) (f / y) - log(f / y) - 1
```

QLIKE loss `(f / y) - log(f / y) - 1`, conditioning variable
one-day-lagged VIX, AIC pre-whitening (`prewhiten = -1`, equivalent
to Ox `PreWhiten = 2`), no trimming, and `R = 10000` bootstrap
replications throughout — matching the call signature in
`Empirics_Volatility.ox`.

## Figure 2 — JNJ, one-versus-one CSPA

```{r jnj-losses}
L_jnj <- sapply(models, function(m) qlike(llq2022_jnj[[m]], llq2022_jnj$rv))
X_jnj <- llq2022_jnj$vix_lag
```

Benchmark is HAR throughout; competitors are AR(1) (left panel) and
HARQ (right panel). A negative loss differential indicates the
benchmark is outperformed; the CSPA test rejects if the dashed
confidence bound dips below zero anywhere over the support of VIX.

```{r fig2-left, fig.width = 6.5, fig.height = 4}
set.seed(20260512)
g_har_ar1 <- cspa_test_plot(
  Y = as.matrix(L_jnj[, "AR1"] - L_jnj[, "HAR"]),
  X = X_jnj, level = 0.05, trim = 0, prewhiten = -1L,
  xlab = "VIX (lagged)", ylab = "QLIKE diff (AR(1) − HAR)"
) + ggplot2::ggtitle("HAR vs AR(1)") +
    ggplot2::geom_hline(yintercept = 0, linewidth = 0.3)
g_har_ar1
```

```{r fig2-right, fig.width = 6.5, fig.height = 4}
set.seed(20260512)
g_har_harq <- cspa_test_plot(
  Y = as.matrix(L_jnj[, "HARQ"] - L_jnj[, "HAR"]),
  X = X_jnj, level = 0.05, trim = 0, prewhiten = -1L,
  xlab = "VIX (lagged)", ylab = "QLIKE diff (HARQ − HAR)"
) + ggplot2::ggtitle("HAR vs HARQ") +
    ggplot2::geom_hline(yintercept = 0, linewidth = 0.3)
g_har_harq
```

```{r fig2-decisions}
fig2 <- function(competitor) {
  Y <- as.matrix(L_jnj[, competitor] - L_jnj[, "HAR"])
  set.seed(20260512)
  r <- cspa_test(Y, X_jnj, level = 0.05, trim = 0, prewhiten = -1L,
                 preselect = TRUE, R = 10000L)
  data.frame(competitor = competitor,
             theta      = unname(r$theta),
             pvalue     = unname(r$pvalue),
             reject     = unname(r$reject))
}
knitr::kable(rbind(fig2("AR1"), fig2("HARQ")), digits = 4, row.names = FALSE)
```

The left panel — AR(1) as the only competitor to HAR — shows the
conditional loss differential staying above zero throughout the VIX
support; the test does not reject. The right panel — HARQ as
competitor — shows the differential dipping clearly below zero in
the VIX ≈ 13-19 range and the confidence bound follows it, so the
CSPA null is rejected. This matches the paper's Figure 2 and the
accompanying text.

## Figure 3 — JNJ, one-versus-all CSPA

Now use all five other models simultaneously as competitors and plot
each \eqn{\hat h_j(x)} (colored), their lower envelope (solid black)
and the upper confidence bound on that envelope (dashed black).
Rejection of the CSPA null occurs when the dashed line is below zero
anywhere.

```{r fig3-ar1, fig.width = 6.5, fig.height = 4.3}
set.seed(20260512)
Y_ar1 <- L_jnj[, setdiff(models, "AR1")] - L_jnj[, "AR1"]
g_ar1 <- cspa_test_plot(
  Y = Y_ar1, X = X_jnj, level = 0.05, trim = 0, prewhiten = -1L,
  xlab = "VIX (lagged)", ylab = "QLIKE diff (competitors − AR(1))"
) + ggplot2::ggtitle("Benchmark: AR(1)") +
    ggplot2::geom_hline(yintercept = 0, linewidth = 0.3)
g_ar1
```

```{r fig3-harq, fig.width = 6.5, fig.height = 4.3}
set.seed(20260512)
Y_harq <- L_jnj[, setdiff(models, "HARQ")] - L_jnj[, "HARQ"]
g_harq <- cspa_test_plot(
  Y = Y_harq, X = X_jnj, level = 0.05, trim = 0, prewhiten = -1L,
  xlab = "VIX (lagged)", ylab = "QLIKE diff (competitors − HARQ)"
) + ggplot2::ggtitle("Benchmark: HARQ") +
    ggplot2::geom_hline(yintercept = 0, linewidth = 0.3)
g_harq
```

```{r fig3-decisions}
fig3 <- function(bench) {
  comp <- setdiff(models, bench)
  Y    <- L_jnj[, comp] - L_jnj[, bench]
  set.seed(20260512)
  r <- cspa_test(Y, X_jnj, level = 0.05, trim = 0, prewhiten = -1L,
                 preselect = TRUE, R = 10000L)
  data.frame(benchmark = bench,
             theta     = unname(r$theta),
             pvalue    = unname(r$pvalue),
             reject    = unname(r$reject))
}
knitr::kable(rbind(fig3("AR1"), fig3("HARQ")), digits = 4, row.names = FALSE)
```

For AR(1) as benchmark the lower envelope sits well below zero
across the VIX support and the dashed bound is below zero in the
low-VIX region — strong CSPA rejection (\eqn{\theta \approx -0.24},
p < 0.001). For HARQ as benchmark the envelope dips modestly below
zero around VIX 25-40 and the dashed bound just barely follows
(\eqn{\theta \approx -0.008}, p ≈ 0.01) — a borderline rejection
at 5%. The paper notes HARQ belongs to the CSMS for 24 of the 28
assets, so for the other four HARQ is itself rejected; on this
dataset, JNJ falls in that small minority.

## Table 4 Panel A — cross-stock rejection counts

For each of 28 stocks, pairwise CSPA tests are run for every
benchmark-competitor pair and rejections are tallied at the 5%
level. Cell \eqn{(k, l)} counts the stocks where the null
"benchmark *l* conditionally dominates alternative *k*" is rejected.
Computed offline by `data-raw/llq2022_uv_cspa.R` (~5 min at
R=10000).

```{r counts-mine}
knitr::kable(llq2022_uv_cspa$mine,
             caption = "forecastdom::cspa_test, R = 10000")
```

```{r counts-paper}
knitr::kable(llq2022_uv_cspa$paper,
             caption = "Published Table_UV_CSPA.xlsx (LLQ 2022)")
```

```{r counts-diff}
diff_mat <- llq2022_uv_cspa$mine - llq2022_uv_cspa$paper
knitr::kable(diff_mat, caption = "Difference (forecastdom − LLQ)")
```

23 of the 30 off-diagonal cells match exactly; 29 are within ±1 and
all 30 within ±2. Residual noise on boundary cells reflects the
different bootstrap seed. The reading is identical to LLQ's:

- **HARQ as benchmark** (column HARQ) — all 28 stocks reject every
  competitor except ARFIMA: HARQ is conditionally superior almost
  everywhere.
- **HARQ / ARFIMA as alternative** (rows HARQ, ARFIMA) — they reject
  every other benchmark.
- **AR(22) as benchmark** (column AR22) — uniformly rejected; the
  simple long-AR is the weakest model.

## S&P 500 — confidence set for the most superior method

The same procedure on the S&P 500 series.

```{r sp500-csms}
L_sp <- sapply(models, function(m) qlike(llq2022[[m]], llq2022$rv))
set.seed(20260512)
cs <- csms(L_sp, llq2022$vix_lag, level = 0.10, trim = 0,
           prewhiten = -1L, preselect = TRUE, R = 10000L,
           method_names = models)
cs
```

The 90% CSMS collapses to `{HARQ}` on S&P 500. ARFIMA is rejected
here even though it survives in 22 of the 28 stocks in the cross-stock
table above — SP500 falls in the reject group.

## Takeaway

Unconditional MSE rankings hide which model is *uniformly* best
across volatility regimes. LLQ's central empirical message — HARQ
and ARFIMA cannot be ruled out as conditionally most superior on
volatility forecasting — is reproduced at three levels: the JNJ
figures (one-vs-one and one-vs-all CSPA), the 28-stock count table,
and the SP500 CSMS.

## References

- Bollerslev, T., Patton, A. J. and Quaedvlieg, R. (2016). Exploiting
  the errors: a simple approach for improved volatility forecasting.
  *Journal of Econometrics*, 192(1), 1-18.
- Corsi, F. (2009). A simple approximate long-memory model of
  realized volatility. *Journal of Financial Econometrics*, 7(2),
  174-196.
- Li, J., Liao, Z. and Quaedvlieg, R. (2022). Conditional Superior
  Predictive Ability. *Review of Economic Studies*, 89(2), 843-875.
