---
title: "Replicating Quaedvlieg (2021)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replicating Quaedvlieg (2021)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 4
)
```

This article demonstrates the two multi-horizon superior predictive
ability tests from Quaedvlieg (2021), implemented in **forecastdom**
as `uspa_mh_test()` and `aspa_mh_test()`. Both tests compare a
*path* of forecasts at horizons \(h = 1, \dots, H\) rather than each
horizon in isolation, and use a moving-block bootstrap to obtain
critical values that account for serial dependence in the
loss-differential path.

* **Uniform SPA** (\(\mathrm{uSPA}\)) — null: the benchmark is at
  least as good as the competitor at *every* horizon. Test statistic
  is the minimum of horizon-wise standardized loss differentials.
* **Average SPA** (\(\mathrm{aSPA}\)) — null: the benchmark is at
  least as good as the competitor on a (user-specified) weighted
  average of horizons. Test statistic is the standardized weighted
  average loss differential.

Throughout, the loss differential is defined as
\(d_{h,t} = L_{\text{bench},h,t} - L_{\text{comp},h,t}\), so positive
entries indicate the benchmark has *higher* loss (is worse) at that
horizon, and rejection of either null requires the benchmark to be
uniformly (or on average) worse.

```{r setup, message = FALSE}
library(forecastdom)
data(quaedvlieg2021)
str(quaedvlieg2021, max.level = 1)
```

The bundled `quaedvlieg2021` object holds the two
loss-differential matrices distributed with the paper's replication
archive: `$uspa` and `$aspa`, each \(1000 \times 20\).

## Per-horizon picture

Before running the tests, it's worth looking at the per-horizon
mean loss differentials. The two example matrices are constructed
so that they discriminate between the tests.

```{r mean-table}
H <- ncol(quaedvlieg2021$uspa)
means <- rbind(
  uspa_dataset = colMeans(quaedvlieg2021$uspa),
  aspa_dataset = colMeans(quaedvlieg2021$aspa)
)
colnames(means) <- paste0("h", seq_len(H))
knitr::kable(round(means, 3))
```

`$uspa` shows positive mean loss differentials at *every* horizon
— the benchmark is worse uniformly. `$aspa` shows positive means
at most horizons but a near-zero mean at the shortest horizon,
so the benchmark is worse on average but not uniformly.

```{r mean-plot, fig.height = 3.5, fig.alt = "Mean loss differential by horizon for the uspa and aspa example matrices"}
library(ggplot2)
df <- data.frame(
  horizon = rep(seq_len(H), 2),
  d_bar   = c(colMeans(quaedvlieg2021$uspa),
              colMeans(quaedvlieg2021$aspa)),
  dataset = rep(c("uspa", "aspa"), each = H)
)
ggplot(df, aes(horizon, d_bar, colour = dataset)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey60") +
  geom_line() + geom_point() +
  labs(x = "Horizon h", y = expression(bar(d)[h]),
       title = "Mean loss differential by horizon") +
  theme_minimal()
```

## Uniform multi-horizon SPA

```{r uspa, results = "hold"}
set.seed(1)
uspa_uspa <- uspa_mh_test(quaedvlieg2021$uspa, L = 3, B = 999)
set.seed(1)
uspa_aspa <- uspa_mh_test(quaedvlieg2021$aspa, L = 3, B = 999)
uspa_uspa
uspa_aspa
```

The uSPA test rejects on `$uspa` (all horizons show the benchmark
losing) but fails to reject on `$aspa` at the 5% level, because
the shortest horizon's mean is essentially zero — the `min` of
standardized horizon-wise means is pulled down by that single
horizon.

## Average multi-horizon SPA

With uniform weights \(w_h = 1/H\) the aSPA statistic is the
standardized unweighted average loss differential:

```{r aspa-uniform, results = "hold"}
w_unif <- rep(1 / H, H)
set.seed(1)
aspa_uspa <- aspa_mh_test(quaedvlieg2021$uspa, weights = w_unif, L = 3, B = 999)
set.seed(1)
aspa_aspa <- aspa_mh_test(quaedvlieg2021$aspa, weights = w_unif, L = 3, B = 999)
aspa_uspa
aspa_aspa
```

The aSPA test rejects on *both* datasets — the negative h=1 mean
of `$aspa` is more than compensated by positive means at longer
horizons, so the weighted average favours rejection. This is the
power gain Quaedvlieg (2021) emphasises: by aggregating across the
forecast path, the average test detects model-level differences
that the per-horizon Diebold-Mariano machinery would miss because
of the multiple-testing burden.

### Custom weights

Down-weighting short horizons makes the aSPA test even more
decisive against `$aspa`:

```{r aspa-down, results = "hold"}
w_down <- c(rep(0, 4), rep(1, 16)) / 16   # zero weight on h = 1..4
set.seed(1)
aspa_mh_test(quaedvlieg2021$aspa, weights = w_down, L = 3, B = 999)
```

Up-weighting the shortest horizons pulls the statistic back toward
zero:

```{r aspa-up, results = "hold"}
w_up <- c(rep(4, 4), rep(0, 16)) / 16     # all weight on h = 1..4
set.seed(1)
aspa_mh_test(quaedvlieg2021$aspa, weights = w_up, L = 3, B = 999)
```

## Block-length sensitivity

The moving-block bootstrap depends on the block length `L`. The
p-value is robust over a reasonable range — small enough to keep
many blocks, large enough to capture the path's serial dependence:

```{r L-sensitivity}
Ls <- c(2, 3, 5, 8, 12)
sens <- do.call(rbind, lapply(Ls, function(L) {
  set.seed(1)
  u <- uspa_mh_test(quaedvlieg2021$uspa, L = L, B = 499)
  set.seed(1)
  a <- aspa_mh_test(quaedvlieg2021$uspa, weights = w_unif, L = L, B = 499)
  data.frame(L = L,
             uspa_stat = u$statistic, uspa_p = u$pvalue,
             aspa_stat = a$statistic, aspa_p = a$pvalue)
}))
knitr::kable(sens, digits = 3, row.names = FALSE)
```

The statistics are independent of `L` (they depend only on the QS
HAC long-run variance, not on the bootstrap), and the p-values are
stable across `L` for this dataset.

## Takeaways

* `uspa_mh_test()` rejects only when the benchmark loses at *every*
  horizon — strong power against uniform underperformance, limited
  power against horizon-specific failures.
* `aspa_mh_test()` rejects when the *weighted-average* differential
  favours the competitor, even if performance at individual horizons
  is mixed. The choice of weights determines which segment of the
  forecast path drives the decision.
* The moving-block bootstrap is essential: forecast paths exhibit
  natural serial dependence, especially when the same model is
  evaluated at successive horizons.

## References

- Andrews, D. W. K. (1991). Heteroskedasticity and autocorrelation
  consistent covariance matrix estimation. *Econometrica*, 59(3),
  817-858.
- Quaedvlieg, R. (2021). Multi-horizon forecast comparison.
  *Journal of Business & Economic Statistics*, 39(1), 40-53.
