---
title: "Replicating Rossi (2006), Table 1: Out-of-Sample Tests"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replicating Rossi (2006), Table 1: Out-of-Sample Tests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 4
)
```

This article replicates the **out-of-sample portion of Table 1** in
Rossi (2006), *"Are exchange rates really random walks? Some evidence
robust to parameter instability"* (*Macroeconomic Dynamics*, 10(1),
20-38), using two functions from **forecastdom**:

* `dm_test()` -- the Diebold-Mariano (1995) test of equal predictive
  accuracy.
* `enc_new()` -- the Clark-McCracken (2001) ENC-NEW encompassing test
  for nested models.

The exercise is the classical Meese-Rogoff (1983) question: at the
monthly horizon, can a small linear AR model of exchange-rate returns
beat a driftless random walk out of sample?

```{r setup, message = FALSE}
library(forecastdom)
library(ggplot2)

data(rossi2006)
str(rossi2006)
```

## The data

Five bilateral nominal exchange rates against the U.S. dollar
(Canada, France, Germany, Italy, Japan), monthly from March 1973 to
December 1998 -- 310 observations per country.

```{r plot-fx}
ggplot(rossi2006, aes(date, log(fx))) +
  geom_line(colour = "#47A5C5") +
  facet_wrap(~ country, scales = "free_y") +
  labs(x = NULL, y = "log(FX)",
       title = "Log nominal exchange rates vs. USD") +
  theme_minimal(base_size = 11)
```

## Forecasting setup

Let \(e_t\) denote the log exchange rate and
\(\Delta e_t = e_t - e_{t-1}\) its monthly return. Two competing
one-step-ahead forecasts of \(\Delta e_{t+1}\):

* **Benchmark -- driftless random walk:**
  \(\widehat{\Delta e}^{RW}_{t+1} = 0\).
* **Alternative -- AR(*p*):** \(\widehat{\Delta e}^{AR}_{t+1}
  = \hat\alpha_t + \sum_{k=1}^{p} \hat\beta_{k,t}\, \Delta e_{t-k+1}\),
  with coefficients re-estimated each period.

Following Rossi, both AR(1) and AR(2) start from the same usable
sample: dropping the first three observations of \(e_t\) gives
\(n_{obs} = T - 3 = 307\) usable monthly returns. The first
\(R = \lceil n_{obs}/2 \rceil = 154\) are used to fit the initial
model and the remaining \(P = 153\) are evaluated out of sample.

The paper considers three estimation schemes:

* **Split (fixed window)** -- coefficients estimated once on \(1:R\)
  and held fixed.
* **Recursive (expanding window)** -- coefficients re-estimated each
  period using all available data.
* **Rolling (fixed-width window)** -- coefficients re-estimated each
  period using the most recent \(R\) observations.

```{r forecast-fn}
forecast_oos <- function(log_fx, p, scheme = c("split", "recursive", "rolling")) {
  scheme <- match.arg(scheme)

  T_full <- length(log_fx)
  dy <- diff(log_fx)

  # Align AR(1) and AR(2) on the same usable sample (drop first 3 obs of e):
  Y  <- dy[3:(T_full - 1)]              # length n_obs = T_full - 3
  L1 <- dy[2:(T_full - 2)]
  L2 <- dy[1:(T_full - 3)]
  Xm <- if (p == 1) matrix(L1, ncol = 1) else cbind(L1, L2)

  n_obs <- length(Y)
  R     <- as.integer(ceiling(n_obs / 2))
  P_oos <- n_obs - R

  e_alt   <- numeric(P_oos)
  e_bench <- numeric(P_oos)

  for (j in seq_len(P_oos)) {
    idx <- switch(scheme,
      split     = seq_len(R),
      recursive = seq_len(R + j - 1),
      rolling   = j:(R + j - 1)
    )
    Z <- cbind(1, Xm[idx, , drop = FALSE])
    b <- as.numeric(solve(crossprod(Z), crossprod(Z, Y[idx])))
    pred <- as.numeric(c(1, Xm[R + j, ]) %*% b)

    e_alt[j]   <- Y[R + j] - pred
    e_bench[j] <- Y[R + j]              # driftless RW: forecast = 0
  }

  list(e_bench = e_bench, e_alt = e_alt, R = R, P = P_oos)
}
```

## Replicating the OOS panel of Table 1

For every combination of country, AR order, and estimation scheme,
we compute:

* The Diebold-Mariano statistic via `dm_test()` with `correction = FALSE`
  to match Rossi's asymptotic \(\chi^2(1)\) reference distribution
  (a two-sided normal in this case).
* The ENC-NEW statistic via `enc_new()`. ENC-NEW critical values are
  non-standard; Clark-McCracken (2001, Table 2) gives 1%, 5%, and
  10% values around 2.65, 1.59, and 0.98 for the AR(1) setup
  (\(k_2-k_1=1,\, \pi=R/n_{obs}\approx 0.5\)) and slightly higher
  for AR(2). We mark the package's statistic at those thresholds.

```{r run-table1}
countries <- levels(rossi2006$country)
schemes   <- c("split", "recursive", "rolling")
orders    <- c(1, 2)

# Approximate Clark-McCracken (2001, Table 2) 5% critical values
# (k1 = 1, k2 - k1 = p, pi = R/n_obs near 0.5)
enc_cv5 <- c(`1` = 1.59, `2` = 2.31)

grid <- expand.grid(
  country = countries,
  p       = orders,
  scheme  = schemes,
  KEEP.OUT.ATTRS = FALSE,
  stringsAsFactors = FALSE
)
grid$DM      <- NA_real_
grid$DM_p    <- NA_real_
grid$ENC     <- NA_real_
grid$ENC_sig <- ""

for (i in seq_len(nrow(grid))) {
  log_fx <- log(subset(rossi2006, country == grid$country[i])$fx)
  fc <- forecast_oos(log_fx, p = grid$p[i], scheme = grid$scheme[i])

  dm  <- dm_test(fc$e_bench, fc$e_alt,
                 alternative = "two.sided", correction = FALSE)
  enc <- enc_new(fc$e_bench, fc$e_alt)

  grid$DM[i]      <- dm$statistic
  grid$DM_p[i]    <- dm$pvalue
  grid$ENC[i]     <- enc$statistic
  grid$ENC_sig[i] <- ifelse(
    enc$statistic > enc_cv5[as.character(grid$p[i])], "*", ""
  )
}
```

### AR(1) results

```{r tab-ar1}
ar1 <- subset(grid, p == 1)
ar1$DM_str  <- sprintf("%.2f (%.2f)", ar1$DM, ar1$DM_p)
ar1$ENC_str <- paste0(sprintf("%.2f", ar1$ENC), ar1$ENC_sig)

dm_ar1 <- reshape(
  ar1[, c("country", "scheme", "DM_str")],
  idvar = "scheme", timevar = "country", direction = "wide"
)
names(dm_ar1) <- gsub("^DM_str\\.", "", names(dm_ar1))

knitr::kable(dm_ar1, row.names = FALSE,
             caption = "DM_T statistic (p-value), AR(1) vs. RW")

enc_ar1 <- reshape(
  ar1[, c("country", "scheme", "ENC_str")],
  idvar = "scheme", timevar = "country", direction = "wide"
)
names(enc_ar1) <- gsub("^ENC_str\\.", "", names(enc_ar1))

knitr::kable(enc_ar1, row.names = FALSE,
             caption = "ENC_T statistic, AR(1) vs. RW. * = exceeds 5% Clark-McCracken critical value")
```

### AR(2) results

```{r tab-ar2}
ar2 <- subset(grid, p == 2)
ar2$DM_str  <- sprintf("%.2f (%.2f)", ar2$DM, ar2$DM_p)
ar2$ENC_str <- paste0(sprintf("%.2f", ar2$ENC), ar2$ENC_sig)

dm_ar2 <- reshape(
  ar2[, c("country", "scheme", "DM_str")],
  idvar = "scheme", timevar = "country", direction = "wide"
)
names(dm_ar2) <- gsub("^DM_str\\.", "", names(dm_ar2))

knitr::kable(dm_ar2, row.names = FALSE,
             caption = "DM_T statistic (p-value), AR(2) vs. RW")

enc_ar2 <- reshape(
  ar2[, c("country", "scheme", "ENC_str")],
  idvar = "scheme", timevar = "country", direction = "wide"
)
names(enc_ar2) <- gsub("^ENC_str\\.", "", names(enc_ar2))

knitr::kable(enc_ar2, row.names = FALSE,
             caption = "ENC_T statistic, AR(2) vs. RW. * = exceeds 5% Clark-McCracken critical value")
```

The DM statistics are typically small in absolute value with
two-sided p-values well above conventional thresholds: the random
walk and the AR are about equally accurate by squared-error loss.
ENC-NEW occasionally exceeds the 5% critical value, especially in
the rolling and recursive schemes -- consistent with Rossi's
finding that small predictability does show up under the
encompassing test once parameter instability is allowed.

## Single-country deep dive: Japan, AR(1), recursive

The print methods give a compact, fully-formatted summary of each
test:

```{r japan}
log_fx_jp <- log(subset(rossi2006, country == "Japan")$fx)
fc_jp     <- forecast_oos(log_fx_jp, p = 1, scheme = "recursive")

dm_test(fc_jp$e_bench, fc_jp$e_alt, alternative = "two.sided",
        correction = FALSE)
enc_new(fc_jp$e_bench, fc_jp$e_alt)
```

The cumulative squared-error differential -- positive when AR(1) is
beating the random walk -- shows that any predictability is
concentrated in narrow sub-samples rather than uniform across the
1986-1998 OOS period:

```{r cum-loss}
loss_diff <- fc_jp$e_bench^2 - fc_jp$e_alt^2
oos_dates <- tail(unique(rossi2006$date), fc_jp$P)

ggplot(data.frame(date = oos_dates, cum = cumsum(loss_diff)),
       aes(date, cum)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line(colour = "#47A5C5", linewidth = 0.8) +
  labs(x = NULL,
       y = "Cumulative SE loss (RW - AR(1))",
       title = "Cumulative squared-error loss differential, Japan",
       subtitle = "Above zero = AR(1) doing better; below = RW doing better") +
  theme_minimal(base_size = 11)
```

## Takeaways

* `dm_test()` reproduces the OOS DM panel of Rossi (2006) Table 1
  with `correction = FALSE` to match the asymptotic \(\chi^2(1)\)
  reference distribution used in the paper.
* `enc_new()` returns the Clark-McCracken statistic; significance
  must be assessed against tabulated non-standard critical values
  (the package documents this in `?enc_new`).
* The "RW is hard to beat" finding emerges clearly in the DM panel,
  while ENC-NEW can pick up the kind of weak, time-varying
  predictability that motivates the parameter-instability tests in
  the rest of the paper.

## References

- Clark, T. E. and McCracken, M. W. (2001). Tests of equal forecast
  accuracy and encompassing for nested models. *Journal of
  Econometrics*, 105(1), 85-110.
- Diebold, F. X. and Mariano, R. S. (1995). Comparing predictive
  accuracy. *Journal of Business & Economic Statistics*, 13(3), 253-263.
- Harvey, D., Leybourne, S., and Newbold, P. (1997). Testing the
  equality of prediction mean squared errors. *International Journal
  of Forecasting*, 13(2), 281-291.
- Meese, R. A. and Rogoff, K. (1983). Empirical exchange rate models
  of the seventies: Do they fit out of sample? *Journal of
  International Economics*, 14(1-2), 3-24.
- Rossi, B. (2006). Are exchange rates really random walks? Some
  evidence robust to parameter instability. *Macroeconomic Dynamics*,
  10(1), 20-38.
