---
title: "Diebold-Mariano on Exchange Rates (Rossi, 2006)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Diebold-Mariano on Exchange Rates (Rossi, 2006)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 4
)
```

This article reproduces the spirit of the out-of-sample test in Rossi (2006),
*"Are exchange rates really random walks? Some evidence robust to parameter
instability"* (*Macroeconomic Dynamics*, 10(1), 20-38), using the
`dm_test()` function from **forecastdom** and the bundled `rossi2006`
dataset.

The exercise asks a classical question from Meese and Rogoff (1983): can a
linear AR model of monthly exchange-rate returns produce out-of-sample
forecasts that beat a driftless random walk?

```{r setup, message = FALSE}
library(forecastdom)
library(ggplot2)

data(rossi2006)
str(rossi2006)
```

## The data

Five bilateral nominal exchange rates against the U.S. dollar
(Canada, France, Germany, Italy, Japan), monthly from March 1973 to
December 1998 — 310 observations per country.

```{r plot-fx}
ggplot(rossi2006, aes(date, log(fx))) +
  geom_line(colour = "#47A5C5") +
  facet_wrap(~ country, scales = "free_y") +
  labs(x = NULL, y = "log(FX)",
       title = "Log nominal exchange rates vs. USD") +
  theme_minimal(base_size = 11)
```

## The forecasting exercise

Let \(e_t\) denote the log exchange rate and \(\Delta e_t = e_t - e_{t-1}\)
its monthly return. Two competing one-step-ahead forecasts of
\(\Delta e_{t+1}\):

* **Benchmark — driftless random walk:** \(\widehat{\Delta e}^{RW}_{t+1} = 0\).
* **Competitor — AR(1):** \(\widehat{\Delta e}^{AR}_{t+1} = \hat\alpha_t + \hat\beta_t \, \Delta e_t\),
  with \((\hat\alpha_t, \hat\beta_t)\) re-estimated by OLS on an
  expanding (recursive) window.

Following Rossi (2006), the first \(R = \lfloor T/2 \rfloor\)
observations are used as the initial estimation window, and the test is
run on the remaining \(P = T - R\) periods.

```{r forecast-fn}
forecast_ar1_vs_rw <- function(log_fx) {
  dlog <- diff(log_fx)
  T_   <- length(dlog)
  R    <- floor(T_ / 2)
  P    <- T_ - R

  e_ar <- numeric(P)
  e_rw <- numeric(P)

  for (j in seq_len(P)) {
    train <- dlog[seq_len(R + j - 1)]
    fit   <- lm(y ~ x, data = data.frame(y = train[-1], x = train[-length(train)]))
    f_ar  <- as.numeric(predict(fit, newdata = data.frame(x = train[length(train)])))
    realised <- dlog[R + j]

    e_ar[j] <- realised - f_ar
    e_rw[j] <- realised - 0
  }

  list(e_ar = e_ar, e_rw = e_rw, P = P)
}
```

## Country-by-country DM test

`dm_test()` takes the two error series and returns the
Diebold-Mariano statistic with the Harvey-Leybourne-Newbold (1997)
small-sample correction by default. The benchmark is the random walk
(`e1`); the competitor is the AR(1) (`e2`).

```{r run-dm}
countries <- levels(rossi2006$country)

dm_results <- lapply(countries, function(co) {
  log_fx <- log(subset(rossi2006, country == co)$fx)
  fc     <- forecast_ar1_vs_rw(log_fx)
  res    <- dm_test(fc$e_rw, fc$e_ar, h = 1, loss = "SE",
                    alternative = "two.sided")
  data.frame(
    country   = co,
    P         = fc$P,
    dm_stat   = unname(res$statistic),
    p_value   = unname(res$pvalue)
  )
})

dm_table <- do.call(rbind, dm_results)
knitr::kable(dm_table, digits = 4,
             caption = "Modified DM test, AR(1) vs. driftless random walk")
```

A negative `dm_stat` would indicate the AR(1) has *lower* squared-error
loss than the random walk; a positive one favours the random walk.
Under the two-sided alternative, p-values close to 1 are consistent
with the Meese-Rogoff result that the random walk is hard to beat at
short horizons.

## A single-country deep dive: Japan

Inspecting one country in more detail. The print method shows the test
configuration and a t-distribution reference (because HLN is on by
default).

```{r japan}
log_fx_jp <- log(subset(rossi2006, country == "Japan")$fx)
fc_jp     <- forecast_ar1_vs_rw(log_fx_jp)
dm_test(fc_jp$e_rw, fc_jp$e_ar, alternative = "two.sided")
```

The cumulative loss differential is a useful diagnostic: it tracks
whether either method dominates persistently or only over short
sub-samples.

```{r cum-loss}
loss_diff <- fc_jp$e_rw^2 - fc_jp$e_ar^2
oos_dates <- tail(unique(rossi2006$date), fc_jp$P)

ggplot(data.frame(date = oos_dates, cum = cumsum(loss_diff)),
       aes(date, cum)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line(colour = "#47A5C5", linewidth = 0.8) +
  labs(x = NULL,
       y = "Cumulative SE loss (RW - AR(1))",
       title = "Cumulative squared-error loss differential, Japan",
       subtitle = "Above zero = AR(1) doing better; below = RW doing better") +
  theme_minimal(base_size = 11)
```

## Takeaways

* `dm_test()` plugs straight into a standard recursive forecasting
  loop and returns the HLN-corrected statistic with t-distribution
  p-values out of the box.
* For these five rates and this sample, the test typically fails to
  reject equal predictive accuracy — consistent with Meese and Rogoff
  (1983) and the broader exchange-rate forecasting literature.
* The cumulative loss-differential plot complements the single number,
  showing *when* the relative performance of the two models shifts —
  the kind of instability that motivates Rossi's parameter-instability
  tests in the original paper.

## References

- Diebold, F. X. and Mariano, R. S. (1995). Comparing predictive
  accuracy. *Journal of Business & Economic Statistics*, 13(3), 253-263.
- Harvey, D., Leybourne, S., and Newbold, P. (1997). Testing the
  equality of prediction mean squared errors. *International Journal
  of Forecasting*, 13(2), 281-291.
- Meese, R. A. and Rogoff, K. (1983). Empirical exchange rate models
  of the seventies: Do they fit out of sample? *Journal of
  International Economics*, 14(1-2), 3-24.
- Rossi, B. (2006). Are exchange rates really random walks? Some
  evidence robust to parameter instability. *Macroeconomic Dynamics*,
  10(1), 20-38.
