---
title: "Replicating Rapach, Strauss & Zhou (2010)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replicating Rapach, Strauss & Zhou (2010)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 4
)
```

This article reproduces the out-of-sample equity premium analysis of
Rapach, Strauss and Zhou (2010, *RFS*) on the bundled `rz2013`
dataset. RSZ test the historical-average benchmark against
predictive regressions on each of 14 Goyal-Welch macro variables
using the Clark-West MSFE-adjusted statistic; we report both
`cw_test()` (their headline statistic) and `enc_new()` (the
Clark-McCracken encompassing test on the same null) side by side.

**Note on the data.** The bundled `rz2013` is the Rapach-Zhou (2013)
*Handbook* vintage of the Welch-Goyal series, not the exact 2008
file RSZ used. R²_OS magnitudes therefore differ in places but the
qualitative pattern of RSZ Table 1 is preserved (interest-rate
predictors significant under CW; valuation ratios not).

```{r setup, message = FALSE}
library(forecastdom)
data(rz2013)

rsz <- subset(rz2013,
              date >= as.Date("1947-01-01") &
              date <= as.Date("2005-12-01"))
```

## Helpers

```{r helper}
recursive_forecasts <- function(y, x, R) {
  P <- length(y) - R
  e1 <- e2 <- f1 <- f2 <- numeric(P)
  for (j in seq_len(P)) {
    ty <- y[1:(R + j - 1)]
    tx <- x[1:(R + j - 1)]
    f1[j] <- mean(ty)
    fit   <- lm(yt ~ xlag,
                data = data.frame(yt   = ty[-1],
                                  xlag = tx[-length(tx)]))
    f2[j] <- as.numeric(predict(fit, newdata = data.frame(xlag = tx[length(tx)])))
    e1[j] <- y[R + j] - f1[j]
    e2[j] <- y[R + j] - f2[j]
  }
  list(e1 = e1, e2 = e2, f1 = f1, f2 = f2)
}
```

## Table 1 — Univariate predictors

Initial estimation 1947:01-1964:12 (R = 216 months); out-of-sample
1965:01-2005:12 (P = 492 months). Columns:

- **R2OS (%)** — out-of-sample R-squared (Campbell-Thompson 2008).
- **CW stat / p-value** — Clark-West MSFE-adjusted test
  (`cw_test()`); RSZ's headline statistic.
- **ENC-NEW** — Clark-McCracken (2001) encompassing test
  (`enc_new()`); non-standard distribution, 5% critical value
  \eqn{\approx 2.7} for \eqn{k_2 = 1} and \eqn{\pi \approx 2.28}.

```{r table1}
predictors <- c("DP", "DY", "EP", "DE", "SVAR", "BM", "NTIS",
                "TBL", "LTY", "LTR", "TMS", "DFY", "DFR", "INFL_lag")

R <- 216

fc_list <- lapply(predictors, function(p) {
  recursive_forecasts(rsz$eq_prem, rsz[[p]], R = R)
})
names(fc_list) <- predictors

tab_uni <- do.call(rbind, lapply(predictors, function(p) {
  fc  <- fc_list[[p]]
  cw  <- cw_test(fc$e1, fc$e2, fc$f1, fc$f2)
  enc <- enc_new(fc$e1, fc$e2)
  data.frame(predictor = p,
             R2OS_pct  = unname(cw$r2os),
             CW_stat   = unname(cw$statistic),
             CW_pvalue = unname(cw$pvalue),
             ENC_NEW   = unname(enc$statistic))
}))

knitr::kable(tab_uni, digits = 3, row.names = FALSE)
```

Predictors with CW p-value < 0.05 — TBL, LTY, LTR, TMS — coincide
with the four interest-rate-related variables RSZ flag as
significant in Table 1. ENC-NEW exceeds the 5% critical value
(~2.7) for the same four, plus DP and DY at the margin.

## Table 2 — Combination forecasts

For each OOS month form a combined forecast across the 14 univariate
regressions; test each combination against the historical average
using both statistics.

```{r table2}
F2    <- sapply(fc_list, function(z) z$f2)
e1    <- fc_list[[1]]$e1
f1    <- fc_list[[1]]$f1
y_oos <- rsz$eq_prem[(R + 1):nrow(rsz)]

combos <- list(
  "Mean"               = rowMeans(F2),
  "Median"             = apply(F2, 1, median),
  "Trimmed mean (10%)" = apply(F2, 1, function(z) mean(z, trim = 0.1))
)

tab_comb <- do.call(rbind, lapply(names(combos), function(nm) {
  f2c <- combos[[nm]]
  e2c <- y_oos - f2c
  cw  <- cw_test(e1, e2c, f1, f2c)
  enc <- enc_new(e1, e2c)
  data.frame(combination = nm,
             R2OS_pct    = unname(cw$r2os),
             CW_stat     = unname(cw$statistic),
             CW_pvalue   = unname(cw$pvalue),
             ENC_NEW     = unname(enc$statistic))
}))

knitr::kable(tab_comb, digits = 3, row.names = FALSE)
```

## Takeaway

Individual macro predictors yield mostly negative R²_OS, in line
with the Welch-Goyal (2008) puzzle. The Clark-West and ENC-NEW
statistics nonetheless reject the historical-average benchmark for
the four interest-rate predictors that RSZ also flag.

Combination forecasts deliver positive R²_OS and reject the
historical-average benchmark under both tests — the central message
of Rapach, Strauss and Zhou (2010): pooling extracts predictive
content that no single predictor reliably captures on its own.

## References

- Clark, T. E. and McCracken, M. W. (2001). Tests of equal forecast
  accuracy and encompassing for nested models. *Journal of
  Econometrics*, 105(1), 85-110.
- Clark, T. E. and West, K. D. (2007). Approximately normal tests
  for equal predictive accuracy in nested models. *Journal of
  Econometrics*, 138(1), 291-311.
- Rapach, D. E., Strauss, J. K. and Zhou, G. (2010). Out-of-sample
  equity premium prediction: combination forecasts and links to the
  real economy. *Review of Financial Studies*, 23(2), 821-862.
- Welch, I. and Goyal, A. (2008). A comprehensive look at the
  empirical performance of equity premium prediction. *Review of
  Financial Studies*, 21(4), 1455-1508.
