---
title: "Replicating Welch & Goyal (2008)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replicating Welch & Goyal (2008)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 4
)
```

This article reproduces the out-of-sample analysis of Welch and
Goyal (2008, *RFS*) for the **log dividend-price ratio** as a
predictor of the annual log equity premium. The bundled `wg2008`
dataset is built from WG's original `PredictorData.xls` (annual
sheet) — the data vintage shipped with the published paper. The
effective sample is **1872-2005 (134 annual observations)**,
matching WG: the file itself begins in 1871, but the one-year lag
in the predictor consumes that row.

The benchmark is the prevailing historical mean (NULL); the
alternative is a predictive regression on the lagged predictor
(ALTERNATIVE). The predictor is constructed as
\eqn{\log(D12_{t-1}) - \log(\text{Index}_{t-1})} per WG Section 1,
where \eqn{D12} is the 12-month moving sum of dividends and
\eqn{\text{Index}} is the S&P 500 price level. (WG's public plotting
script `goyal-welch-plots.R` uses the level D/P instead, but their
paper text and Table 1 use the log form.)

WG report five OOS statistics per predictor in Table 1; this article
computes each on the same data:

- **R²_OS** — out-of-sample R² (Campbell-Thompson 2008).
- **ΔRMSE** — \eqn{\sqrt{\text{MSE}_N} - \sqrt{\text{MSE}_A}}.
- **MSE-F** — McCracken (2007) F-statistic for equal MSE
  (`mse_f_test()`).
- **ENC-NEW** — Clark-McCracken (2001) encompassing test
  (`enc_new()`).
- **CW MSFE-adj** — Clark-West (2007) MSFE-adjusted t-statistic
  (`cw_test()`), reported by WG in footnote 2.

```{r setup, message = FALSE}
library(forecastdom)
library(ggplot2)
data(wg2008)
# WG (2008) Table 1 covers 1872-2005 — the entire bundled file.
wg <- wg2008
c(first_year = min(wg$year), last_year = max(wg$year), n = nrow(wg))
```

## Helper: recursive forecasts (WG procedure)

The recursive setup matches WG's `goyal-welch-plots.R` exactly:

- At each year `t`, refit `lm(logeqp ~ log_dp_lag)` on years 1...t-1.
- ALTERNATIVE forecast for year t = fitted value at the
  contemporaneous `log_dp_lag[t]`.
- NULL forecast = mean of `logeqp` over years 1...t-1.

```{r helper}
recursive_forecasts <- function(y, x, R) {
  n <- length(y)
  P <- n - R
  e_N <- e_A <- f_N <- f_A <- numeric(P)
  for (j in seq_len(P)) {
    idx <- seq_len(R + j - 1)
    f_N[j] <- mean(y[idx])
    fit    <- lm.fit(cbind(1, x[idx]), y[idx])
    f_A[j] <- sum(coef(fit) * c(1, x[R + j]))
    e_N[j] <- y[R + j] - f_N[j]
    e_A[j] <- y[R + j] - f_A[j]
  }
  list(e_N = e_N, e_A = e_A, f_N = f_N, f_A = f_A,
       year = wg$year[(R + 1):n])
}
```

## Table 1 — five tests across three OOS specifications

WG explore three OOS-start dates: 20 years after the data begins
(≈ 1892), 1965, and the most recent 30 years (1976-2005). The
column `R2_bar_OS` is the *adjusted* out-of-sample R² that WG
report in Table 1, applied to the OOS sample of size T:
\eqn{\bar R^2_{OS} = 1 - (1 - R^2_{OS})\,(T-1)/(T-k)} with k = 2
parameters (intercept + predictor).

```{r main-table}
specs <- list(
  list(label = "20 yr after start (1892+)", R = 20L),
  list(label = "1965 onward",              R = which(wg$year == 1964)),
  list(label = "1976 onward (recent 30)",  R = which(wg$year == 1975))
)

run_spec <- function(spec) {
  fc    <- recursive_forecasts(wg$logeqp, wg$log_dp_lag, R = spec$R)
  MSE_N <- mean(fc$e_N^2)
  MSE_A <- mean(fc$e_A^2)
  T_oos <- length(fc$e_N)
  R2    <- 1 - MSE_A / MSE_N
  R2bar <- 1 - (1 - R2) * (T_oos - 1) / (T_oos - 2)
  dRMSE <- sqrt(MSE_N) - sqrt(MSE_A)
  msef  <- mse_f_test(fc$e_N, fc$e_A)
  enc   <- enc_new(fc$e_N, fc$e_A)
  cw    <- cw_test(fc$e_N, fc$e_A, fc$f_N, fc$f_A)
  data.frame(spec         = spec$label,
             R_est        = spec$R,
             T_oos        = T_oos,
             R2_OS_pct    = 100 * R2,
             R2_bar_OS    = 100 * R2bar,
             dRMSE_pct    = 100 * dRMSE,
             MSE_F        = unname(msef$statistic),
             ENC_NEW      = unname(enc$statistic),
             CW_stat      = unname(cw$statistic),
             CW_p         = unname(cw$pvalue))
}
tab <- do.call(rbind, lapply(specs, run_spec))
knitr::kable(tab, digits = 3, row.names = FALSE)
```

Compared to WG's reported d/p numbers:

| Spec | T_oos | This article \eqn{\bar R^2_{OS}} | WG \eqn{\bar R^2_{OS}} | Gap | Source in WG paper |
|---|---:|---:|---:|---:|---|
| 1892+ (20 yr after start) | 114 | **-2.06** | **-2.06** | 0.00 | p. 1474 in-text table ("All years") |
| 1965+ | 41 | -3.73 | -3.69 | 0.04 | Table 1, "Forecasts begin 1965" column |
| Recent 30 yr (1976+) | 30 | -15.22 | -15.14 | 0.09 | p. 1474 in-text table ("Recent 30 years") |

The longest window matches WG **exactly to two decimals**. The two
shorter windows are within 0.1 percentage points. The pattern of
the residual gap — zero on the long sample, small on the short
samples that are entirely post-1965 — is consistent with minor
revisions to Goyal's annual data file between the 2007 vintage
that fed the published paper and the version currently distributed
through Goyal's website. The longest window draws most of its
weight from pre-1965 data that hasn't been revised; the shorter
windows are entirely post-1965 and show small drift in proportion
to how concentrated they are. Other plausible reconstructions
(WG's `log(1 + R − Rfree)` plotting-script formula, hybrid
Shiller/CRSP returns, alternative OOS-start boundaries) do not
close the gap.

The deterioration of DP through time is unmistakable: a small
negative R²_OS in the long sample, deeper-negative in 1965+, and
substantially negative for 1976-2005. McCracken (2004) and
Clark-McCracken (2001) asymptotic 5% critical values for k₂ = 1
extra regressor:

| π = P/R | MSE-F (5%) | ENC-NEW (5%) |
|---:|---:|---:|
| 0.6  | 1.62 | 2.37 |
| 1.0  | 1.71 | 2.52 |
| 2.0  | 1.82 | 2.70 |

For the 1892+ window (π ≈ 6.4) MSE-F is small and ENC-NEW is below
the 5% threshold; for the recent 30 years both statistics are
firmly negative or near zero. No window supports a "DP beats the
mean" conclusion under WG's tests.

## WG Figure 1 — cumulative SSE difference for DP

WG's signature visual is the cumulative squared-error difference
\eqn{\sum_{s \le t}(e_{N,s}^2 - e_{A,s}^2)}: a rising line means
the ALTERNATIVE beats the NULL up to that date, a falling line
the opposite. The plot below mirrors the *d/p* panel of WG Figure 1
(IS = dotted, OOS = solid; Oil Shock 1973-1975 shaded in red).

```{r wg-fig1, fig.width = 7.5, fig.height = 4.5}
# IS residuals from a single regression on the entire sample
fit_full <- lm(logeqp ~ log_dp_lag, data = wg)
is_xy    <- residuals(fit_full)
is_mean  <- wg$logeqp - mean(wg$logeqp)

# OOS residuals starting at year 21
R <- 20L
fc <- recursive_forecasts(wg$logeqp, wg$log_dp_lag, R = R)

is_imp  <- cumsum(is_mean^2) - cumsum(is_xy^2)
oos_imp <- c(rep(NA, R), cumsum(fc$e_N^2) - cumsum(fc$e_A^2))

df <- data.frame(year = wg$year, IS = is_imp, OOS = oos_imp)
df_long <- rbind(
  data.frame(year = df$year, kind = "IS",  value = df$IS),
  data.frame(year = df$year, kind = "OOS", value = df$OOS)
)

ggplot(df_long, aes(x = year, y = value, color = kind,
                    linetype = kind)) +
  annotate("rect",
           xmin = 1973, xmax = 1975,
           ymin = -Inf, ymax = Inf,
           fill = "red", alpha = 0.15) +
  geom_hline(yintercept = 0, linewidth = 0.3) +
  geom_line(linewidth = 0.9, na.rm = TRUE) +
  scale_color_manual(values = c(IS = "black", OOS = "steelblue4")) +
  scale_linetype_manual(values = c(IS = "dashed", OOS = "solid")) +
  labs(x = NULL,
       y = "Cumulative SSE difference (NULL − ALT)",
       title = sprintf("d/p, 1872 - %d", max(wg$year))) +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.position = "top")
```

The replicated picture matches the *d/p* panel of WG Figure 1: a
quiet first half-century, a climb from WW II to the early 1970s
where DP modestly beats the historical mean, a peak around the Oil
Shock, and a steep decline through the 1990s as the dividend
yield collapsed during the dot-com bull market. The IS line (dashed)
sits steadily above zero — DP looks like a useful in-sample
predictor — while the OOS line (solid) eventually crashes through
zero in the late 1990s, the gap between IS and OOS that motivated
WG's "comprehensive look".

## Takeaway

For the dividend-price ratio at annual frequency, applied with WG's
own data and procedure:

- **Long sample (1892+)** — every OOS statistic agrees that DP does
  not significantly beat the historical mean; the encompassing
  evidence is weak.
- **1965+ window** — both R²_OS and MSE-F turn clearly negative.
- **Recent 30 years (1976+)** — the DP-augmented forecast is
  decisively worse than the historical mean.

This is precisely WG's central message about the dividend-price
ratio: in-sample significance does not survive an honest
out-of-sample evaluation, and the cumulative-SSE plot makes the
structural break around the Oil Shock and the late-1990s decline
immediate.

## References

- Clark, T. E. and McCracken, M. W. (2001). Tests of equal forecast
  accuracy and encompassing for nested models. *Journal of
  Econometrics*, 105(1), 85-110.
- Clark, T. E. and West, K. D. (2007). Approximately normal tests
  for equal predictive accuracy in nested models. *Journal of
  Econometrics*, 138(1), 291-311.
- McCracken, M. W. (2007). Asymptotics for out of sample tests of
  Granger causality. *Journal of Econometrics*, 140(2), 719-752.
- Welch, I. and Goyal, A. (2008). A comprehensive look at the
  empirical performance of equity premium prediction. *Review of
  Financial Studies*, 21(4), 1455-1508.
